---
---
---

# Final Project

```{r}
library(arrow)
library(tidyverse)
library(lubridate)

data_path <- "/Users/aoluwolerotimi/Datasets/" # AO Changed file path
applications <- read_feather(paste0(data_path,"app_data_starter_coded.feather"))
```

## clean the columns

```{r}
columns_with_x_suffix <- grep("\\.x$", names(applications), value = TRUE)

names(applications)[names(applications) %in% columns_with_x_suffix] <- sub("\\.x$", '', columns_with_x_suffix)


applications <- applications %>% select(-ends_with(".y"))
```

## Feature Engineering

```{r}
#quarter 
applications <- applications %>%
  mutate(
    quarter = paste0(year(filing_date), "/", quarter(filing_date)),
  )

# Aggregate applications by quarter and examiner
applications <- applications %>%
  group_by(quarter, examiner_id) %>%
  mutate(new_applications = n_distinct(application_number)) %>%
  ungroup()

applications <- applications %>%
  group_by(quarter, examiner_id) %>%
  mutate(ISSUED_applications = sum(disposal_type == "ISS" & !duplicated(application_number)))

applications <- applications %>%
  group_by(quarter, examiner_id) %>%
  mutate(abn_applications = sum(disposal_type == "ABN" & !duplicated(application_number)))

applications <- applications %>%
  group_by(quarter, examiner_id) %>%
  mutate(PEN_applications = sum(disposal_type == "PEND" & !duplicated(application_number)))

applications <- applications %>%
  group_by(quarter,examiner_art_unit) %>%
  mutate(examiner_art_unit_num =  n_distinct(examiner_id))%>%
  ungroup()
```

```{r}
max_quarter <- "2017/1"

applications <- applications %>%
        filter(quarter <= max_quarter)

# separation_indicator
applications <- applications %>%
  group_by(examiner_id) %>%
  mutate(max_quarter_examiner = max(quarter)) %>%
  ungroup() %>%
  mutate(separation_indicator = if_else(max_quarter_examiner < max_quarter, 1, 0))

# to delete lastest_date that is beyond the max_quarter
#max_date <- "2017-03-31"
#applications <- applications %>%
#  filter(latest_date <= as.Date(max_date))
```

```{r}
#au_move_indicator
applications <- applications %>%
  group_by(examiner_id) %>%
  mutate(au_move_indicator = if_else(examiner_art_unit != lag(examiner_art_unit), 1, 0)) %>%
  ungroup()

applications <- applications %>%
  mutate(au_move_indicator = if_else(is.na(au_move_indicator), 0, au_move_indicator))

#tc_move_indicator
applications <- applications %>%
  group_by(examiner_id) %>%
  mutate(tc_move_indicator = if_else(tc != lag(tc), 1, 0)) %>%
  ungroup()

applications <- applications %>%
  mutate(tc_move_indicator = if_else(is.na(tc_move_indicator), 0, tc_move_indicator))
```

```{r}
# get the count of au_moves by quarter
applications <- applications %>%
  group_by(examiner_id, quarter) %>%
  mutate(
    au_moves = sum(au_move_indicator)
  ) %>%
  ungroup()

#tc_moves by quarter
applications <- applications %>%
  group_by(examiner_id, quarter) %>%
  mutate(
    tc_moves = sum(tc_move_indicator)
  ) %>%
  ungroup()

#distinct uspc_class and uspc_subclass by quarter
applications <- applications %>%
  group_by(examiner_id, quarter) %>%
  mutate(
    num_classes = n_distinct(uspc_class),
    num_subclasses = n_distinct(uspc_subclass)
  ) %>%
  ungroup()
```

```{r}
# get the process time
applications <- applications %>%
  mutate(
    patent_issue_date = as.Date(patent_issue_date),
    abandon_date = as.Date(abandon_date),
    processing_time = case_when(
      disposal_type == "ISS" ~ as.numeric(patent_issue_date - filing_date, units = "days"),
      disposal_type == "ABN" ~ as.numeric(abandon_date - filing_date, units = "days"),
      TRUE ~ 0
    )
  )

applications <- applications %>%
  mutate(
    iss_time = ifelse(disposal_type == "ISS", processing_time, NA),
    abn_time = ifelse(disposal_type == "ABN", processing_time, NA)
  )

# Computing averages within the same dataframe
applications <- applications %>%
  group_by(examiner_id) %>%
  mutate(
    avg_processing = mean(processing_time, na.rm = TRUE),
    avg_ISS_processing = mean(iss_time, na.rm = TRUE),
    avg_ABN_processing = mean(abn_time, na.rm = TRUE)
  ) %>%
  ungroup()
```

## Covariates Cleaning

```{r}
columns_to_exclude <- c(
  "examiner_art_unit", "examiner_art_unit_num",
  "women_in_art_unit"
) # Due to a high number of missing vlaue in the gender value, the quality of the women_in_art_unit is poor, thus excluded

df <- applications[, !(names(applications) %in% columns_to_exclude)]
colSums(is.na(df))
```

```{r}
#drop the na examiner_id rows
df <- subset(df, !is.na(examiner_id))

#aggregate to quarter
quarter_df <- df %>%
  group_by(examiner_id) %>%
  distinct(quarter, .keep_all = TRUE) %>%
  select(examiner_id, quarter, latest_date, separation_indicator, ISSUED_applications, PEN_applications, abn_applications, gender, race, tenure_days, au_moves, tc_moves, num_classes, num_subclasses, avg_processing, avg_ISS_processing, avg_ABN_processing) %>%
  arrange(examiner_id, quarter)
```

```{r}
#collapse to individual observation
collapsed_df <- quarter_df %>%
  group_by(examiner_id) %>%
  summarize(
    gender = first(gender),
    race = first(race),
    tenure_days = first(tenure_days),
    ISSUED_applications = sum(ISSUED_applications),
    abandoned_applications = sum(abn_applications),
    au_moves = sum(au_moves),
    tc_moves = sum(tc_moves),
    PEN_applications = sum(PEN_applications) / n(),
    separation_indicator = max(separation_indicator),
    num_classes = sum(num_classes),
    num_subclasses = sum(num_subclasses),
    avg_processing = first(avg_processing),
    avg_ISS_processing = first(avg_ISS_processing),
    avg_ABN_processing = first(avg_ABN_processing)
  )

#append NA with 'unknown'
collapsed_df <- collapsed_df %>%
  mutate(gender = ifelse(is.na(gender), "unknown", gender))
```

```{r}
# saving the feather file
# write_feather(collapsed_df, "/Users/aoluwolerotimi/Datasets/collapsed_df_final.feather")
```

```{r}
colSums(is.na(collapsed_df))
```

```{r}
#Fill NA to 0 
collapsed_df <- collapsed_df %>%
  mutate_all(~ifelse(is.na(.), 0, .))
```

```{r}
summary(collapsed_df)
```

## Model Building

```{r}
collapsed_df$separation_indicator <- factor(make.names(as.character(collapsed_df$separation_indicator)))

collapsed_df$gender <- as.factor(collapsed_df$gender)
collapsed_df$race <- as.factor(collapsed_df$race)

#Train Split
library(caret)

set.seed(123) 
splitIndex <- createDataPartition(collapsed_df$separation_indicator, p = .75, list = FALSE)
train_data <- collapsed_df[splitIndex,]
test_data <- collapsed_df[-splitIndex,]

indexes <- sample(1:nrow(collapsed_df), size = 0.75 * nrow(collapsed_df))
train_data_lm <- collapsed_df[indexes, ]
test_data_lm <- collapsed_df[-indexes, ]
```

-   **What are the organizational and social factors associated with the length of patent application prosecution?**

    Hypothesis - The impact of tenure on outcomes like ISSUED applications, abandoned applications, or average processing times might differ between genders/races

```{r}
train_control <- trainControl(method = "cv", number = 10)

#avg_processing as outcome
model <- train(avg_processing ~ tenure_days + num_classes + num_subclasses +
               ISSUED_applications + au_moves + tc_moves + gender + race +
               gender * tenure_days + race * tenure_days,
               data = train_data_lm,
               method = "lm", 
               trControl = train_control)

#avg_ISS_processing as outcome
model2 <- train(avg_ISS_processing ~ tenure_days + num_classes + num_subclasses+
               ISSUED_applications + au_moves + tc_moves + gender + race +
               gender * tenure_days + race * tenure_days,
               data = train_data_lm,
               method = "lm", 
               trControl = train_control)

#avg_ABN_processing as outcome
model3 <- train(avg_ABN_processing ~ tenure_days + num_classes + num_subclasses+
               ISSUED_applications + au_moves + tc_moves + gender + race +
               gender * tenure_days + race * tenure_days,
               data = train_data_lm,
               method = "lm", 
               trControl = train_control)
```

```{r}
summary(model)
```

```{r}
summary(model2)
```

```{r}
summary(model3)
```

Insights :

-   **What are the organizational and social factors associated with examiner attrition**

```{r}
train_control <- trainControl(method = "cv", number = 10, classProbs = TRUE, summaryFunction = twoClassSummary)

attrition_model <- train(separation_indicator ~ tenure_days + ISSUED_applications + abandoned_applications + avg_processing +
                         avg_ISS_processing + avg_ABN_processing + au_moves + tc_moves +
                         gender + race + gender * tenure_days + race * tenure_days,
                         data = train_data,
                         method = "glm",
                         family = "binomial",
                         trControl = train_control,
                         preProcess = "scale", 
                         metric = "ROC") 
```

```{r}
summary(attrition_model)
```

insights:
